Automated Lensing Learner - I: An automated Strong Lensing Identification Pipeline

### 2017-03-09 ms.tex
update results to Creg

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:pipeline}
\begin{center}
\includegraphics[scale=0.55]{figures/supervised_classification_workflow.png}
\caption{** Cartoon of pipeline ** from
  http://www.nltk.org/book/ch06.html}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In this section, we discuss the Receiver Operating Characteristic
(ROC) curve, which shows the true positive rate (tpr) as a function of
false positive rate (fpr) for a given model and test set.  The true
positive rate is defined as the number of true positives divided by
the total number of positives (completeness of the positive
identifications).  The false positive rate is defined as the number of
false positives divided by the total number of negatives (impurity of
the identifications).  The ROC curve illustrates the performance of
our trained model as we vary the discrimination threshold.

@@

% HST model performance
Figure~\ref{fig:ROC} shows the ROC curves for three different HOG
parameterizations for the HST-like dataset in red, green, and blue.
The parameters and corresponding AUC for each curve are labeled in the
legend.  The models for the HST-like dataset perform well, and have
AUCs very close to 1.

% LSST model performance
On the other hand, the model for our LSST-like dataset (solid magenta)
does not perform well.  The dashed diagonal like corresponds to a
model that is as good as a random classifier, e.g. a coin-flip, for
each test image.  The LSST-like ROC curve does marginally better, and
the $AUC=0.527$ is within the statistical noise of the sample.

% No BCG LSST
One possible method of improving the detection of strong lensing
systems is to model out the BCG.  The dotted magenta ROC curve
corresponds to a model trained and tested on LSST-like data where the
BCG has not been added to the mock image.  This model results in an
AUC of 0.64, performing better than random.  In other words, if we
were to perfectly model out the BCG contribution in LSST-like data, we
could improve the performance by $\sim10\%$.

% Choosing C
The ROCs correspond to models that have been trained with logistic
regression regularization parameter, $C_\text{LogReg}$, that best
describes the training dataset without overfitting.  We discuss the
choice of regularization parameter in
Section~\ref{sec:regularizedfit}.

%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
%% \begin{figure}[t]\label{fig:ROCHST}
%% \begin{center}
%% \includegraphics[width=\columnwidth]{figures/ROC_HST_HOG.png}
%% \caption{Red, green, blue: ROC curves for a models trained on the
%%   HST-like data with the Logistic Regression algorithm.  The magenta
%%   solid (dotted) curve show the ROC curve for the LSST-like data with
%%   (without) the BCG.  The corresponding parameters are labeled: N for
%%   number of orientations, ppc for pixels-per-cell, cpb for
%%   cells-per-block, and C for best regularization parameter.  The AUC
%%   for each model is also labeled in the legend, where an AUC close to
%%   1 is a near-ideal model, and an AUC close to 0.5 is a model that
%%   performs as well as a coin-flip.}
%% \end{center}
%% \end{figure}
%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:ROCcompilation}
\begin{center}
\includegraphics[width=\columnwidth]{figures/metrics/ROC/ROC_compilation.pdf}
\caption{Red, blue, green: ROC curves for models trained on our whole
  10,000 training set and tested on our holdout set of 1,000.  These
  respectively correspond to the HST, LSST 1 year, and LSST 10 year
  data.  The solid lines are for data that include the lensing central
  galaxy, and the dashed lines for the data where there is no lensing
  galaxy, mimicking an ideal removal of the lens.  Model performance
  can be summarized by the area under the curve (AUC), labeled in the
  legend. $AUC=1$ is a perfect model, and $AUC=0.5$ is a useless
  model.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:PRcompilation}
\begin{center}
\includegraphics[width=\columnwidth]{figures/metrics/PR/PR_compilation.pdf}
\caption{Red, blue, green: Precision-recall (PR) curves for models
  trained on our whole 10,000 training set and tested on our holdout
  set of 1,000.  These respectively correspond to the same models and
  data shown in Figure~\ref{fig:ROCcompilation}.  An ideal model would
  reach both a precision (purity) and recall (completeness) that equal
  1.  Note, that this performance describes a data set with a 50-50
  split between lens and non-lens containing images.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Logistic regression and regularization
As described in Section~\ref{sec:LR}, LR trains a model with
complexity determined by the regularization parameter coefficient,
$C_\text{LogReg}$.  Increasing model complexity (larger
$C_\text{LogReg}$), will better describe features in the training
set[wc?].  However, an overly complex model will overfit the training
set at the expense of its performance on any independent test set.
The regularization parameter ultimately defines the model performance,
and we must perform a parameter search to identify the optimal value
for $C_\text{LogReg}$.

%  Our running over C - score and AUC as characterizations
Figure~\ref{fig:regularization} shows the model performance as a
function of regularization parameter.  The left panel shows the
accuracy, or score, of the model as a function of regularization
parameter.  The accuracy is simply the percentage of correct
predictions, true positives and true negatives, out of a given set.
The right panel shows the AUC of the model as a function of
regularization parameter.  The solid and dotted lines respectively
correspond to the model performance on the test and training set.

@@

% Train time with Creg
Lower Creg, more regularized, smaller volume to search for the
solution (best fit coefficients), so faster on average.  But, still
has some randomness associated with the path it takes to convergence
(not purely monotonic).

% Scores vs. regularization on larger train/test set with comparable
% grid search in parameterization
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure*}[t]\label{fig:regularization}
\begin{center}
\includegraphics[width=.66\columnwidth]{figures/Creg/HST.pdf}
\includegraphics[width=.66\columnwidth]{figures/Creg/LSST1.pdf}
\includegraphics[width=.66\columnwidth]{figures/Creg/LSST10.pdf}
\caption{AUC of the model with varying LR regularization coefficient
  parameter, $C_{reg}$, used when training the model classifier.  We
  use a subset of the the 10,000 training images to search over the LR
  $C_{reg}$ parameter, training on 8,000 and testing on 1,000.  Each
  panel corresponds to a different mock observation.  From left to
  right: HST, LSST for one year, and LSST for 10 years.  The xsolid
  blue lines correspond to the AUC of the training set, and the dotted
  blue lines to the AUC of the test set.  To avoid overfitting, we
  choose the smallest parameter for which the AUC of the test set is
  maximal: 5000, 10, and 5000, respectively.  In thin red solid lines,
  we show the train time of the model, which roughly increases in a
  log-log scaling with logistic regression coefficient parameter.}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Scores vs. regularization on larger train/test set with comparable
% grid search in parameterization
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:trainsizeLSST10}
\begin{center}
\includegraphics[width=\columnwidth]{figures/trainsize/LSST10.pdf}
\caption{Solid blue line: AUC for models with varying size of the
  training set for LSST10.  Dotted red line: Train time for that
  model.  The improvement of AUC scales linearly with the log of the
  training set size, but train time roughly scales logarithmically.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:trainsizenLSST10}
\begin{center}
\includegraphics[width=\columnwidth]{figures/trainsize/nLSST10.pdf}
\caption{Same as Figure~\ref{fig:trainsizeLSST10} for nLSST10.  For
  small trainsize, the nLSST10 performs better than LSST10. }
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:pipeline}
\begin{center}
\includegraphics[scale=0.55]{figures/supervised_classification_workflow.png}
\caption{Cartoon of pipeline \footnote{from
    http://www.nltk.org/book/ch06.html} for supervised classification.
  In our case, the labels are either lens or non-lens containing
  images.  Our input is the set of mock HST- or LSST-like
  observations.  The feature extractor is the histogram of oriented
  gradients, producing an N-dimensional feature vector that quantify
  edges in our images.  Our machine learning algorithm is logistic
  regression.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In this section, we discuss the Receiver Operating Characteristic
(ROC) curve (see Figure~\ref{fig:ROCcompilation}), which shows the
true positive rate (tpr) as a function of false positive rate (fpr)
for a given model and test set.  The true positive rate is defined as
the number of true positives divided by the total number of positives
(completeness of the positive identifications).  The false positive
rate is defined as the number of false positives divided by the total
number of negatives (impurity of the identifications).  The ROC curve
illustrates the performance of our trained model as we vary the
discrimination threshold.

@@

% Precision-recall
For comparison, we also show a Precision-Recall curve (see
Figure~\ref{fig:PRcompilation}), where precision is the number of true
positives identified as positive divided by the total number of
positively identified images (purity) and recall is the number of true
positives identified as positive divided by the total number of true
positives (completeness).  Each point in the figure is calculated with
a varying threshold for identification.  Since our sample has a class
balance of 50-50 between positives (lens) and negatives (non-lens),
the most lenient threshold for identification would yield a precision
of 0.5 at a recall of 1.0.  It is important to note that this figure
changes as the class balance changes; if we had 90\% non-lenses and
10\% lenses, the most lenient threshold would yield a precision of 1/9
at a recall of 1.  The ROC curve is a more typical metric for
supervised classification.

% Summary of ROC curves
Figure~\ref{fig:ROCcompilation} shows the ROC curves for models
trained using the entire 10,000 training sample, with best-case HOG
and regularization parameters.  The models have been evaluated on a
hold-out test set of 1,000 images that were not used in the parameter
search.  We show the mock HST, LSST1, and LSST10 results respectively
in red, blue, and green.  Solid lines correspond to a model trained
and tested on images with the lensing galaxy.  Dashed lines correspond
to a model trained and tested on images where the lensing galaxy is
excluded from the mock observation, simulating an ideal modeling and
subtraction of the lensing galaxy, which has been one proposed method
to improve the identification of strong lensing systems.  The
corresponding AUC is listed in the legend.

% Model performance summary
The model performance for the mock HST data is AUC=0.975 for images
with the lens galaxy, and AUC=0.98 for images without the lens galaxy
(red solid and dashed).  On the other hand, the model for our
LSST-like dataset for one year has an AUC=0.625 with the lens galaxy
and AUC=0.579 without the lens galaxy (blue solid and dashed), and the
model for our LSST-like dataset for 10 years has an AUC=0.809 with the
lens galaxy and AUC=0.792 without the lens galaxy (green solid and
dashed).  Removal of the lens galaxy does not systematically perform
better, and is actually dependent on the size of the training set.  We
discuss relative model performance and complexity for images with and
without the lens galaxy in Section~\ref{sec:trainsetsize}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:ROCcompilation}
\begin{center}
\includegraphics[width=\columnwidth]{figures/metrics/ROC/ROC_compilation.pdf}
\caption{Red, blue, green: ROC curves for models trained on our whole
  10,000 training set and tested on our holdout set of 1,000.  These
  respectively correspond to the HST, LSST 1 year, and LSST 10 year
  data.  The solid lines are for data that include the lensing central
  galaxy, and the dashed lines for the data where there is no lensing
  galaxy, mimicking an ideal removal of the lens.  Model performance
  can be summarized by the area under the curve (AUC), labeled in the
  legend. $AUC=1$ is a perfect model, and $AUC=0.5$ is a useless
  model.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%            
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:PRcompilation}
\begin{center}
\includegraphics[width=\columnwidth]{figures/metrics/PR/PR_compilation.pdf}
\caption{Red, blue, green: Precision-recall (PR) curves for models
  trained on our whole 10,000 training set and tested on our holdout
  set of 1,000.  These respectively correspond to the same models and
  data shown in Figure~\ref{fig:ROCcompilation}.  An ideal model would
  reach both a precision (purity) and recall (completeness) that equal
  1.  Note, that this performance describes a data set with a 50-50
  split between lens and non-lens containing images.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Logistic regression and regularization
As described in Section~\ref{sec:LR}, LR trains a model with
complexity determined by the regularization parameter coefficient,
$C_\text{LogReg}$.  Increasing model complexity (larger
$C_\text{LogReg}$), will better describe features in the training set.
However, an overly complex model will overfit the training set at the
expense of its performance on any independent test set.  The
regularization parameter ultimately defines the model performance, and
we must perform a parameter search to identify the optimal value for
$C_\text{LogReg}$.

%  Our running over C - score and AUC as characterizations
Figure~\ref{fig:regularization} shows the model performance as a
function of regularization parameter for each data set HST, LSST, and
LSST10.  The solid and dotted lines respectively correspond to the
model performance on the test and training set.

@@

% Train time with Creg
In red, we show the train time as a function of $C_\text{LogReg}$.
Lower Creg, more regularized, smaller volume to search for the
solution (best fit coefficients), so faster on average.  But, still
has some randomness associated with the path it takes to convergence
(not purely monotonic).

% Scores vs. regularization on larger train/test set with comparable
% grid search in parameterization
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure*}[t]\label{fig:regularization}
\begin{center}
\includegraphics[width=.66\columnwidth]{figures/Creg/HST.pdf}
\includegraphics[width=.66\columnwidth]{figures/Creg/LSST1.pdf}
\includegraphics[width=.66\columnwidth]{figures/Creg/LSST10.pdf}
\caption{AUC of the model with varying LR regularization coefficient
  parameter, $C_{reg}$, used when training the model classifier.  We
  use a subset of the the 10,000 training images to search over the LR
  $C_{reg}$ parameter, training on 8,000 and testing on 1,000.  Each
  panel corresponds to a different mock observation.  From left to
  right: HST, LSST for one year, and LSST for 10 years.  The solid
  blue lines correspond to the AUC of the test set, and the dotted
  blue lines to the AUC of the training set.  To avoid overfitting, we
  choose the smallest parameter for which the AUC of the test set is
  maximal: 5000, 10, and 5000, respectively.  In thin red solid lines,
  we show the train time of the model, which roughly increases in a
  log-log scaling with logistic regression coefficient parameter.}
\end{center}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Scores vs. regularization on larger train/test set with comparable
% grid search in parameterization
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                        
\begin{figure}[t]\label{fig:trainsizeLSST10}
\begin{center}
\includegraphics[width=\columnwidth]{figures/trainsize/LSST10_with_nLSST10.pdf}
\caption{Solid (dotted) blue line: AUC for models with varying size of
  the training set for LSST10 (nLSST10).  Solid (dotted) red line:
  Optimal regularization parameter for training on the train size for
  LSST10 (nLSST10). The improvement of AUC scales linearly with the
  log of the training set size, but train time roughly scales
  logarithmically.  Note, LSST10 requires more model complexity to
  exceed the performance of nLSST10.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
